---
title: "VBPQR_TrendReport"
author: 
  name: "John Ryan Kivela, MA"
  email: "Ryan.Kivela@narbha.org"
  affiliation: "The Alliance ACO"
editor: visual
date: today
date-format: long
format:
  html:
    theme: pulse
    embed-resources: true
    toc: true
    toc-depth: 6
    code-fold: false
    footnotes-hover: true
---

# Abstract

This report aggregates all VBP QR files into one master file that will be used to assess baseline and performance levels across measures and providers.

```{r}
#| label: Setup
#| include: false
#| echo: false
#| warning: false
#| error: false

# Load Libraries
library(tidyverse)
library(purrr)
library(readxl)

# Create provider shortnames based on TIN
# Create a data frame called ProviderShortnames
ProviderShortnames <- 
  data.frame(
  TIN = c("86-0206928", "86-0207499", "86-0214457", "86-0215065", "86-0223720", "86-0250938", "86-0290033", "86-0473221", "94-2880847"),
  ProviderShortname = c("PH", "SHG", "MMHC", "CPIH", "TGC", "LCBHC", "SBHS", "EHS", "CBI")
  )
```

# Data Discovery

This data is collected from the BCBSAZ Health Choice health plan on a monthly basis. A separate file is received per provider, per month. The source and quality of this data is already well documented in the VBP Validation for Follow Up After Hospitalization and Anti-depressant Medication Management, so we will not re-iterate it here.

Prior to importing to the model, each VBP Quality Roster data file is copied to a Master Folder containing all Quality Roster reports for all months of the measurement year, for all of the providers.

## Import Data files

These are excel files and require importing of a specific sheet. The name of the sheet has varied in name and location within the file, sometimes appearing a literal sheet, and sometimes appearing only in the data model view. 

```{r}
# Function to read Excel files from a folder and its subfolders
read_excel_sheets_as_separate_df <- function(path) {
  # Get list of files in the current directory
  files <- list.files(path = path, full.names = TRUE, recursive = TRUE)
  
  # Initialize an empty list to store data frames
  all_dfs <- list()
  
  # Loop through each Excel file
  for (file in files) {
    if (grepl("\\.xlsx$|\\.xls$", file, ignore.case = TRUE)) {
      file_name <- basename(file)
      sheet_names <- excel_sheets(file)
      
      # Read each sheet and store as a separate data frame
      for (sheet in sheet_names) {
        df_name <- paste0(file_name, "_", sheet)
        data <- readxl::read_excel(file, sheet = sheet)
        assign(df_name, data)
        all_dfs[[df_name]] <- get(df_name)
      }
    }
  }
  
  return(all_dfs)
}

# Replace './data/Master' with your folder path
folder_path <- "./data/VBPReports/Master/2022"

# Call the function to read Excel files and sheets as separate data frames
result_separate_df <- read_excel_sheets_as_separate_df(folder_path)

# Filter list items that end in "_Detail"
filtered_list <- result_separate_df[sapply(names(result_separate_df), function(x) grepl("_Detail$", x))]

# Bind tibbles/data frames into one single data frame
combined_df <- bind_rows(filtered_list)

# Filter out rows with NA in the second column
filtered_combined_df <- 
  combined_df |> 
  filter(!is.na(`...2`))

# Extract values from the first row of filtered_combined_df
first_row_values <- unlist(filtered_combined_df[1,])

# Store the values in a list called DetailSheetNames
DetailSheetNames <- as.list(first_row_values)

# Remove rows where the value in the first column is "LOB"
filtered_combined_df <- 
  filtered_combined_df |> 
  filter(`Health Choice Arizona (HCA) – Arizona Medicaid` != "LOB")

# Convert DetailSheetNames to character vector
DetailSheetNames <- as.character(DetailSheetNames)

# Rename columns using values from DetailSheetNames
names(filtered_combined_df) <- DetailSheetNames

# Merge filtered_combined_df with ProviderShortnames based on TIN column
merged_df <- merge(filtered_combined_df, ProviderShortnames, by.x = "Health Home TIN", by.y = "TIN", all.x = TRUE)

# Convert excel dates to a yyyy-mm-dd format
merged_df$`Report Period` <- as.Date(as.numeric(merged_df$`Report Period`), origin = "1899-12-30")
merged_df$`Data Period` <- as.Date(as.numeric(merged_df$`Data Period`), origin = "1899-12-30")
merged_df$`Measurement End Date` <- as.Date(as.numeric(merged_df$`Measurement End Date`), origin = "1899-12-30")
merged_df$CapPeriodDateFormat <- as.Date(as.numeric(merged_df$CapPeriodDateFormat), origin = "1899-12-30")

# Rename columns with 2023 column names
merged_df <-
  merged_df |> 
  rename(`Gap Status` = Numerator)

# Update numerator values (0,1) to Gap Status values (Open, CLosed)
merged_df <- merged_df %>%
  mutate(`Gap Status` = ifelse(`Gap Status` == 0, "OPEN", "CLOSED"))

# Split Member Name into LastName and FirstName
merged_df <- merged_df %>%
  separate(`Member Name`, into = c("LastName", "FirstName"), sep = ",")

# Select variables for analysis
Data_VBPQR_2022_AllRecordsCombined <-
  merged_df |> 
  select(
    `Report Period`,
    `Data Period`,
    `Measurement End Date`,
    `Health Home TIN`,
    ProviderShortname,
    `Health Home Name`,
    `Member ID`,
    LastName,
    FirstName,
    `Measure ID`,
    `SubMeasure ID`,
    `SubMeasure Description`,
    `Gap Status`)
```












