---
title: "VBPQR_TrendReport"
author: 
  name: "John Ryan Kivela, MA"
  email: "Ryan.Kivela@narbha.org"
  affiliation: "The Alliance ACO"
editor: visual
date: today
date-format: long
format:
  html:
    theme: pulse
    embed-resources: true
    toc: true
    toc-depth: 6
    code-fold: false
    footnotes-hover: true
---

# Abstract

This report aggregates all VBP QR files into one master file that will be used to assess baseline and performance levels across measures and providers.

# Setup

```{r}
#| label: Setup
#| include: false
#| echo: false
#| warning: false
#| error: false

# Load Libraries
library(tidyverse)
library(purrr)
library(readxl)
library(stringr)
library(scales)
library(lubridate)

# Create provider shortnames based on TIN
# Create a data frame called ProviderShortnames
ProviderShortnames <- 
  data.frame(
  TIN = c("86-0206928", "86-0207499", "86-0214457", "86-0215065", "86-0223720", "86-0250938", "86-0290033", "86-0473221", "94-2880847"),
  ProviderShortname = c("PH", "SHG", "MMHC", "CPIH", "TGC", "LCBHC", "SBHS", "EHS", "CBI")
  )

# Create a list with the ultimate column names for the data
VBPTrendReport_colnames <- 
  c("ClaimsAdjudicatedThrough",
    "AllianceProviderTIN",
    "ProviderShortname",
    "HealthHomeName",
    "MemberID",
    "LastName",
    "FirstName",
    "MeasureID",
    "Measure",
    "GapStatus")

ContractMeasures <- c("AMM", "FUH", "FUM", "HDO", "FUA")

```

# Data Discovery

This data is collected from the BCBSAZ Health Choice health plan on a monthly basis. A separate file is received per provider, per month. The source and quality of this data is already well documented in the VBP Validation for Follow Up After Hospitalization and Anti-depressant Medication Management, so we will not re-iterate it here. ***add references here***

Prior to importing to the model, each VBP Quality Roster data file is copied to a Master Folder containing all Quality Roster reports for all months of the measurement year, for all of the providers.

These are excel workbooks and contain multiple sheets of data, of which we need only specific sheets. The names and format of these sheets has varied in composition, name, and location within the file, sometimes appearing as a literal sheet, and sometimes appearing only in the data model view. In addition, the variables' format changed from 2022-2023, requiring a reworking of the data wrangle code. Thus this model is built in two parts, one for 2022, and one for 2023 and on.

# Import Measurement Year 2022

## Procedure

```{r}
# Create Function to read Excel files from a folder and its subfolders
read_excel_sheets_as_separate_df <- function(path) {
  # Get list of files in the current directory
  files <- list.files(path = path, full.names = TRUE, recursive = TRUE)
  
  # Initialize an empty list to store data frames
  all_dfs <- list()
  
  # Loop through each Excel file
  for (file in files) {
    if (grepl("\\.xlsx$|\\.xls$", file, ignore.case = TRUE)) {
      file_name <- basename(file)
      sheet_names <- excel_sheets(file)
      
      # Read each sheet and store as a separate data frame
      for (sheet in sheet_names) {
        df_name <- paste0(file_name, "_", sheet)
        data <- readxl::read_excel(file, sheet = sheet)
        assign(df_name, data)
        all_dfs[[df_name]] <- get(df_name)
      }
    }
  }
  
  return(all_dfs)
}

# Set the folder path
folder_path <- "./data/VBPReports/Master"

# Call the function to read Excel files and sheets as separate data frames
result_separate_df <- read_excel_sheets_as_separate_df(folder_path)

# This result is a list of tibbles that are each individual sheet from the workbook, stored individually from one another. 

# Filter list items that end in "_Detail". This isolates the "Detail" sheet which contains data on each case, including its compliance status.
filtered_list <- result_separate_df[sapply(names(result_separate_df), function(x) grepl("_Detail$", x))]

# Bind tibbles/data frames into one single data frame that contains all of the data from each of the Detail sheets
combined_df <- bind_rows(filtered_list)

# Clean and polish data

# The next few steps process each sheet to remove superfluous rows that provide narrative text when viewed in excel, but are not needed for pure data.

# Filter out rows with NA in the second column.
filtered_combined_df <- 
  combined_df |> 
  filter(!is.na(`...2`))

# Extract values from the first row of filtered_combined_df to be used as the column headers
first_row_values <- unlist(filtered_combined_df[1,])

# Store the values in a list called DetailSheetNames
DetailSheetNames <- as.list(first_row_values)

# Convert DetailSheetNames to character vector
DetailSheetNames <- as.character(DetailSheetNames)

# Remove rows where the value in the first column is "LOB". This removes the column headers that was imported from the excel. Several of these column names come in as uninformative data, like "...2".
filtered_combined_df <- 
  filtered_combined_df |> 
  filter(`Health Choice Arizona (HCA) – Arizona Medicaid` != "LOB")

# Rename columns using values from DetailSheetNames
names(filtered_combined_df) <- DetailSheetNames

# Merge filtered_combined_df with ProviderShortnames based on TIN column
merged_df <- merge(filtered_combined_df, ProviderShortnames, by.x = "Health Home TIN", by.y = "TIN", all.x = TRUE)

# Convert excel dates to a yyyy-mm-dd format
merged_df$`Report Period` <- as.Date(as.numeric(merged_df$`Report Period`), origin = "1899-12-30")
merged_df$`Data Period` <- as.Date(as.numeric(merged_df$`Data Period`), origin = "1899-12-30")
merged_df$`Measurement End Date` <- as.Date(as.numeric(merged_df$`Measurement End Date`), origin = "1899-12-30")
merged_df$CapPeriodDateFormat <- as.Date(as.numeric(merged_df$CapPeriodDateFormat), origin = "1899-12-30")

# Filter for contract measures
merged_df <- 
  merged_df |> 
  filter(`Measure ID` %in% ContractMeasures)


# # Update numerator values (0,1) to Gap Status values (Open, Closed). In the 2022 data, HDO data is reverse coded. This is corrected here so that the measures can be viewed consistently.

# Assuming merged_df is your dataframe

merged_df <- merged_df |>
  mutate(
    `Gap Status` = if_else(`Measure ID` == "AMM" & Numerator == 1, "CLOSED",
                           if_else(`Measure ID` == "FUH" & Numerator == 1, "CLOSED",
                                   if_else(`Measure ID` == "FUM" & Numerator == 1, "CLOSED",
                                           if_else(`Measure ID` == "HDO" & Numerator == 1, "OPEN", 
                                                   if_else(`Measure ID` == "HDO" & Numerator == 0, "CLOSED", "OPEN"))))))
      
      


# Split Member Name into LastName and FirstName
merged_df <- merged_df |>
  separate(`Member Name`, into = c("LastName", "FirstName"), sep = ",")

# Select variables for analysis
Data_VBPQR_2023_Detail <-
  merged_df |> 
  select(
    `Data Period`,
    `Health Home TIN`,
    ProviderShortname,
    `Health Home Name`,
    `Member ID`,
    LastName,
    FirstName,
    `Measure ID`,
    `SubMeasure Description`,
    `Gap Status`)

# Assuming df is your data frame and VBPTrendReport_colnames is your list of column names
colnames(Data_VBPQR_2023_Detail) <- VBPTrendReport_colnames

```

# Import Measurement Year 2023 and on

## Procedure

The format and structure of the VBP Quality Roster was changed at the beginning of the 2023 Measurement Year. In particular, the variable "Numerator" with binary values (0 and 1) representing compliance was changed to "Gap Status" with values of "OPEN" or "CLOSED". Thus, we have a separate procedure for importing data for the measurement year 2023, and presumable on into the future. Also, in the 2023 format, the "Controls", including report dates, are stored on a separate sheet from the "Roster" which has the case details and compliance status. These 2 sheets must be individually imported, processed, and bound before further analysis is possible.

```{r}
# Set the folder path
folder_path <- "./data/VBPReports/Master"

# Find files containing "Roster" in their filenames
roster_files <- list.files(path = folder_path, pattern = "Roster", full.names = TRUE)

# Function to read "Roster" and "Controls" sheets from a file and bind them
bind_roster_controls <- function(file) {
  # Read sheets "Roster" and "Controls"
  roster_sheet <- tryCatch(readxl::read_excel(file, sheet = "Roster"), error = function(e) NULL)
  controls_sheet <- tryCatch(readxl::read_excel(file, sheet = "Controls"), error = function(e) NULL)
  
  # Rename "Health Home TIN & Name" column to "Assigned Health Home TIN & Name" in Controls sheet
  col_index <- which(names(controls_sheet) == "Health Home TIN & Name")
  if (length(col_index) > 0) {
    names(controls_sheet)[col_index] <- "Assigned Health Home TIN & Name"
  }
  
  # Check if both sheets are not NULL and have data
  if (!is.null(roster_sheet) && !is.null(controls_sheet) && !all(is.na(roster_sheet)) && !all(is.na(controls_sheet))) {
    # Add "Report TIN" column to Roster sheet
    roster_sheet$`Report TIN` <- substr(basename(file), 1, 10)
    
    # Bind sheets and fill missing spaces with NA
    max_rows <- max(nrow(roster_sheet), nrow(controls_sheet))
    roster_filled <- bind_rows(roster_sheet, data.frame(matrix(NA, nrow = max_rows - nrow(roster_sheet), ncol = ncol(roster_sheet))))
    controls_filled <- bind_rows(controls_sheet, data.frame(matrix(NA, nrow = max_rows - nrow(controls_sheet), ncol = ncol(controls_sheet))))
    
    # Combine "Roster" and "Controls" sheets
    combined_data <- cbind(roster_filled, controls_filled)
    
    return(combined_data)
  } else {
    message("One or both sheets ('Roster' and/or 'Controls') not found or contain no data in file:", file)
    return(NULL)
  }
}

# Create an empty list to store combined data from all files
combined_data_list <- list()

# Iterate through each file in roster_files and store it in "Combined_data_list"
for (file in roster_files) {
  combined_data <- bind_roster_controls(file)
  
  # Add combined data to the list
  if (!is.null(combined_data)) {
    combined_data_list[[file]] <- combined_data
  }
}

# Iterate through combined_data_list and update columns 19-24 to maintain consistent variable names and populate the Report Date for all cases.
for (file in names(combined_data_list)) {
  df <- combined_data_list[[file]]
  
  # Check if columns 19-24 have a non-NA first value and fill the entire column
  for (col in 19:24) {
    if (!is.na(df[1, col])) {
      df[, col] <- df[1, col]
    }
  }
  
  # Update the list with modified dataframe
  combined_data_list[[file]] <- df
}

# Combining all rows from all tibbles in combined_data_list
all_combined_data <- bind_rows(combined_data_list)

# Fill down the values in the 'CapMonth' column
all_combined_data <- all_combined_data %>%
  fill(CapMonth, .direction = "down")


# the result of this is a data frame that is the Roster sheet, combined with the Control sheet, for each individual work book, bound together for all of the months for all of the providers.

# Clean and polish the data

# Remove rows with NA in column "...2"
filtered_all_combined_data <-
  all_combined_data |> 
  filter(!is.na(...2))

# Extracting column names from the first row for columns 1-8
col_names <- as.character(filtered_all_combined_data[1, 1:9])  # Extract column names from the first row for columns 1-18

# Assigning the first row as column headers for columns 1-18 and removing the first row from these columns
names(filtered_all_combined_data)[1:9] <- col_names  # Assign column names for columns 1-9
# Remove rows where the column "Measure" has the value "Measure". This removes the original column names.
filtered_all_combined_data <- filtered_all_combined_data[filtered_all_combined_data$Measure != "Measure", ]

# Change 'CapMonth' to the last day of the prior month
filtered_all_combined_data <- filtered_all_combined_data %>%
  mutate(CapMonth = lubridate::floor_date(CapMonth, "month") - 1,
         CapMonth = format(CapMonth, "%Y-%m-%d"))

# Replace the claims adjudicated date with the capmonth date
filtered_all_combined_data <-
  filtered_all_combined_data |> 
  mutate(ClaimsAdjudicatedDate = CapMonth)

# Select Variables for analysis
filtered_all_combined_data <-
  filtered_all_combined_data |> 
  select(
    Measure,
    `Gap Status`,
    `Member ID`,
    `Member Age`,
    `Member Name`,
    `Health Home TIN & Name`,
    `Assigned Health Home TIN & Name`,
    `Report TIN`,
    MeasurementYearEndDate,
    ClaimsAdjudicatedDate,
    CapMonth,
    LastRefresh
  )

# split member name into first name/lastname
filtered_all_combined_data <- filtered_all_combined_data |>
  separate(`Member Name`, into = c("LastName", "FirstName"), sep = ",", remove = FALSE)  # Splitting "Member Name" column into LastName and FirstName

# Split Health Home TIN into separate columns for each
filtered_all_combined_data <- filtered_all_combined_data |>
  separate(`Health Home TIN & Name`, into = c("Health Home TIN", "HealthHomeName"), sep = " ", extra = "merge")  # Splitting column into Health Home TIN and HealthHomeName

# Trim the health home name for extra spaces and characters
filtered_all_combined_data$HealthHomeName <- substr(filtered_all_combined_data$HealthHomeName, start = 3, stop = nchar(filtered_all_combined_data$HealthHomeName))

# Merge filtered_combined_df with ProviderShortnames based on TIN column
filtered_all_combined_data <- merge(filtered_all_combined_data, ProviderShortnames, by.x = "Report TIN", by.y = "TIN", all.x = TRUE)

# Extract the first three digits from "Measure" column into a new column "Measure ID"
filtered_all_combined_data$`Measure ID` <- substr(as.character(filtered_all_combined_data$Measure), 1, 3)

# Remove the first 4 digits from the 'Measure' column
filtered_all_combined_data$Measure <- substr(filtered_all_combined_data$Measure, start = 7, stop = nchar(filtered_all_combined_data$Measure))

# Remove the first 3 digits from the 'Member ID' column
filtered_all_combined_data$`Member ID` <- substring(filtered_all_combined_data$`Member ID`, 4)

# Select Variables for analysis

Data_VBPQR_2023_Roster <-
  filtered_all_combined_data |> 
  select(
    ClaimsAdjudicatedDate,
    `Report TIN`,
    ProviderShortname,
    HealthHomeName,
    `Member ID`,
    LastName,
    FirstName,
    `Measure ID`,
    Measure,
    `Gap Status`
  )

# Assuming df is your data frame and VBPTrendReport_colnames is your list of column names
colnames(Data_VBPQR_2023_Roster) <- VBPTrendReport_colnames


```

# Finalize data set

The 2022 and 2023 data sets are now aggregated, cleaned, and formatted so that they can be combined with one another. This step binds the 2 data frames into one. This step also offers the opportunity to clean up the environment from transitional data that is no longer needed.

```{r}

# Bind the 2022 data to the 2023 and on data
Data_VBPQR_AllRecords <-
  rbind(Data_VBPQR_2023_Detail,
        Data_VBPQR_2023_Roster)

# Clean up the environment from all of the transitional data frames

# List all objects in the environment
objects_to_keep <- c("Data_VBPQR_AllRecords")  # Add more objects to keep as needed

# Get a list of objects in the environment
all_objects <- ls()

# Remove objects except for those specified in objects_to_keep
objects_to_remove <- all_objects[!all_objects %in% objects_to_keep]
#rm(list = objects_to_remove)

```

# Process the data set

Now that all of the data is formatted and combined, we will need to trim it down to contract measures only.

```{r}

# Write to CSV
write.csv(Data_VBPQR_AllRecords, "./data/output/Data_VBPQR_AllRecords.csv")

# Run power automate to copy this file to Alliance Intranet Support folder
```

# Begin the analysis

```{r}

# Creating a table that counts occurrences of "OPEN" and "CLOSED" per group
counts_table <- 
  Data_VBPQR_AllRecords |>
  group_by(ClaimsAdjudicatedThrough, ProviderShortname, MeasureID) |>
  summarise(
    Open_Count = sum(GapStatus == "OPEN", na.rm = TRUE),
    Closed_Count = sum(GapStatus == "CLOSED", na.rm = TRUE),
    Total_Count = Open_Count + Closed_Count
  ) |>
  mutate(
    Percentage_Closed = Closed_Count / Total_Count, accuracy = 0.01
  ) |> 
  mutate(Target = if_else(MeasureID == "AMM", .5729,
                          if_else(MeasureID == "FUH", .3900,
                                  if_else(MeasureID == "FUM", .5750,
                                          if_else(MeasureID == "HDO", .93,
                                                  0))))) |> 
  mutate(HPBenchmark = if_else(MeasureID == "AMM", .6040,
                          if_else(MeasureID == "FUH", .4600,
                                  if_else(MeasureID == "FUM", .5040,
                                          if_else(MeasureID == "HDO", .93,
                                                  0))))) |> 
  mutate(Performance_Target = Percentage_Closed - Target) |> 
  mutate(Performance_HPBenchmark = Percentage_Closed - HPBenchmark)



# Write to csv
write.csv(counts_table, "./data/output/VBPQR_Alliance_AllMeasures.csv")

```

```{r}


```
